---
title: "Prediction of Patient Status using Voice Measurements"
author: "INSOFE Lab Assignment on Logistic Regression - Batch 30"
date: "8 July 2017"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
---

**NOTE** Before starting this assignment please remember to clear your environment, you can do that by running the following code chunk

```{r}
rm(list = ls(all = TRUE))

```

# Agenda 

* Get the data

* Data Pre-processing

* Build a model

* Predictions

* Communication


# Reading & Understanding the Data

* The "parkinsons_data.csv"" dataset is composed of a range of biomedical voice measurements. Each column in the table is a particular voice measure, and each row corresponds one of 195 voice recordings.

* The dataset has 195 rows and 23 columns. 

* The column/variable names' explanation is given below:

1) __MDVP:Fo(Hz) :__ Average vocal fundamental frequency

2) __MDVP:Fhi(Hz) :__ Maximum vocal fundamental frequency

3) __MDVP:Flo(Hz) :__ Minimum vocal fundamental frequency

4) __MDVP:Jitter(%); MDVP:Jitter(Abs); MDVP:RAP; MDVP:PPQ; Jitter:DDP :___ Several measures of variation in fundamental frequency

5) __MDVP:Shimmer; MDVP:Shimmer(dB); Shimmer:APQ3; Shimmer:APQ5; MDVP:APQ; Shimmer:DDA :__ Several measures of variation in amplitude

6) __NHR; HNR :__ Two measures of ratio of noise to tonal components in the voice

7) __status :__ Health status of the subject (Parkinson's) - Positive, (Normal) - Healthy

8) __RPDE; D2 :__  Two nonlinear dynamical complexity measures

9) __DFA :__ Signal fractal scaling exponent

10) __spread1; spread2; PPE :__ Three nonlinear measures of fundamental frequency variation

* Make sure the dataset is located in your current working directory and read in the data

```{r}
setwd("I:/DATA-SCIENCE/Insofe/Assignments/LogisticRegression/Logistics Assignment")
parkinsons.data = read.csv("parkinsons_data.csv", header = TRUE)
```

* Use the str() function to get a feel for the dataset.

```{r}
str(parkinsons.data)
summary(parkinsons.data)
```


* Take a look at the data using the "head()" and "tail()" functions

```{r}
head(parkinsons.data)
tail(parkinsons.data)

```

* Are there any missing values in the dataset?

```{r}
sum(is.na(parkinsons.data))

```

# Data Pre-processing

## Train/Test Split

* Split the data 70/30 into train and test sets, using __Stratified Sampling__ by setting the seed as "786"

```{r}
library(caret)
set.seed(786)
trainrows = createDataPartition(parkinsons.data$status, p = 0.7, list = FALSE) 
train.data = parkinsons.data[trainrows,]
test.data = parkinsons.data[-trainrows,]


```

# Build a model

## Basic Logistic Regression Model

* Use the glm() function to build a basic model

* Build a model using all the variables, excluding the response variable, in the dataset

```{r}
library(nnet)
model.complete.variables = multinom(formula = status ~., data = train.data)
model.complete.variables = glm(formula = status ~., family = "binomial", data = train.data)

```

* Get the summary of the model and understand the output

```{r}
summary(model.complete.variables)
tail(train.data, 20)
```

## Create an ROC Curve

1) Get a list of predictions (probability scores) using the predict() function

```{r}

probability.train = predict(model.complete.variables, type = "response")

```


2) Using the ROCR package create a "prediction()" object

```{r}
library(ROCR)
predicted.values = prediction(probability.train, train.data$status)

```


3) Extract performance measures (True Positive Rate and False Positive Rate) using the "performance()" function from the ROCR package

```{r}
performance = performance(predicted.values, measure = "tpr", x.measure = "fpr")

```


4) Plot the ROC curve using the extracted performance measures (TPR and FPR)

```{r}
plot(performance, col = rainbow(10) , colorize = T, print.cutoffs.at=seq(0,1,0.1) )

```

* Extract the AUC score of the ROC curve and store it in a variable named "auc"

```{r}

auc.performance = performance(predicted.values, measure = "auc")
auc.performance@y.values[[1]]

```


## Choose a Cutoff Value

* Based on the trade off between TPR and FPR depending on the business domain, a call on the cutoff has to be made.

```{r}

### Write your answer here

# A cutoff of ----- seems reasonable
# If we are more concerned about the specificity, the cut-off rate should be of a higher value (~ 0.8).
# If we need to keep a balance between specificity and sensitivity, the cut-ff rate should be in balance (~0.7)
```


## Predictions on test data

* After choosing a cutoff value, predict the class labels on the test data using our model

```{r}

prob.test = predict(model.complete.variables, test.data, type = "response")
predict.test = ifelse(prob.test > 0.7, "Parkinson's", "Normal")

```

# Evaluation Metrics for classification

## Manual Computation

### Confusion Matrix

* Create a confusion matrix using the table() function

```{r}

conf.matrix = table(test.data$status, predict.test)


```

### Specificity

* Calculate the Specificity

* The Proportion of correctly identified negatives by the test/model.

$${Specificity} = \frac{Number~of~True~Negatives}{Number~of~True~Negatives + Number~of~False~Positives}$$

```{r}
specificity = conf.matrix[1,1]/sum(conf.matrix[1,])


```


### Sensitivity

* Calculate the Sensitivity

* The Proportion of correctly identified positives by the test/model.

$${Sensitivity} = \frac{Number~of~True~Positives}{Number~of~True~Positives + Number~of~False~Negatives}$$

```{r}

sensitivity = conf.matrix[2,2]/sum(conf.matrix[2,])

```

### Accuracy

* Calculate the Accuracy

* The Proportion of correctly identified psotivies/negatives in the entire population by the test/model

$${Accuracy} = \frac{Number~of~True~Positives +Number~of~True~Negatives}{Number~Of~Subjects~in~the~Population}$$

```{r}

accuracy = sum(diag(conf.matrix))/sum(conf.matrix)

```

## Automated Computation through Caret

* Use the caret package to compute the evaluation metrics

* Mention your reference positive class as an argument to the confusionMatrix() function

```{r}

library(caret)
confusionMatrix(predict.test, test.data$status, positive = "Parkinson's")

```











